---
title: "PML"
output: html_document
---
## Overview

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

## Loading Libraries
```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(corrplot)
library(e1071)
```

## Data Loading and Processing
```{r}
train <- read.csv('./pml-training.csv', header=T)
valid <- read.csv('./pml-testing.csv', header=T)
train$classe<-as.factor(train$classe)
dim(train)
```

## Cleaning Input Data

```{r}
trainData<- train[, colSums(is.na(train)) == 0]
validData <- valid[, colSums(is.na(valid)) == 0]
dim(trainData)
```
```{r}
dim(validData)
```
```{r}
trainData <- trainData[, -c(1:7)]
validData <- validData[, -c(1:7)]
```

## Preparing Dataset
```{r }
set.seed(1234) 
library(caret)
inTrain <- createDataPartition(trainData$classe, p = 0.7, list = FALSE)
trainData <- trainData[inTrain, ]
testData <- trainData[-inTrain, ]

```

## Cleaning even further by removing the variables that are near-zero-variance
```{r}
NZV <- nearZeroVar(trainData)
trainData <- trainData[, -NZV]
testData  <- testData[, -NZV]
dim(trainData)
```

## Correlation Matrix
```{r}
cor_mat <- cor(trainData[, -53])
corrplot(cor_mat, order = "FPC", method = "color", type = "upper", 
         tl.cex = 0.8, tl.col = rgb(0, 0, 0))
```

## Print highly correlated
```{r}
highlyCorrelated = findCorrelation(cor_mat, cutoff=0.75)
names(trainData)[highlyCorrelated]

```

## Model building
We will use
1) Decision Trees
2)Random Forest

## Prediction with classification Trees

We will first use this method and also visualize it to better understand the algorithm
```{r}
set.seed(12345)
decisionTreeMod1 <- rpart(classe ~ ., data=trainData, method="class")
fancyRpartPlot(decisionTreeMod1)
```
We will then validate the model on the testData
```{r}
predictTreeMod1 <- predict(decisionTreeMod1, testData, type = "class")
cmtree <- confusionMatrix(predictTreeMod1, testData$classe)
cmtree
```
## Plot
To get a better understanding of our results, we will visualize it:
```{r}
plot(cmtree$table, col = cmtree$byClass, 
     main = paste("Decision Tree - Accuracy =", round(cmtree$overall['Accuracy'], 4)))
```

## Random Forest
Now we will try another model, Random Forest. It is an ensemble machine learning algorithm. It is a combination of trees and the result is chosen by voting

```{r}
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modRF1 <- train(classe ~ ., data=trainData, method="rf", trControl=controlRF)
modRF1$finalModel
```

We validate the model now
```{r}
predictRF1 <- predict(modRF1, newdata=testData)
cmrf <- confusionMatrix(predictRF1, testData$classe)
cmrf
```
The accuracy rate using the random forest is very high: Accuracy : 1 and therefore the out-of-sample-error is equal to 0***. But it might be due to overfitting.
Let's plot the model

```{r}
plot(modRF1)

```

```{r}
plot(cmrf$table, col = cmrf$byClass, main = paste("Random Forest Confusion Matrix: Accuracy =", round(cmrf$overall['Accuracy'], 4)))

```


The best model is Random Forest hence we will use it on the validation data
```{r}
Results <- predict(modRF1, newdata=validData)
Results

```
